\subsection{SDMetrics}
SDMetrics is a very complete design measurement tool, analysing a wide range of UML diagrams, including Class, Usecase, Activity and Statemachine diagrams. For each type of diagram, this tool generates several metrics.\\
For example, \textbf{NumAttr} metric, one of the metrics that has already been adressed in this paper, is one of the metrics that is calculated from Class diagrams. Other one is the \textbf{ExtPts} metric, wich is calculated from Usecase diagrams, and gives us the number of extension points of a given use case.\\
SDMetrics is written in java, and provides us a graphical user interface. The type of source files it receives to analyse are \textit{XMI} files. \textit{XMI} (\textbf{X}ML \textbf{M}etadata \textbf{I}nterchange)  is an \textit{OMG} (Object Management Group) standard to generate XML-based representations of UML and other OO data models. Most modeling tools support project export in XMI.\\
After analysing the different metrics for the project this tool allows us to access the results from different views. We will approach the ones that seem the most important:
\begin{itemize}
\item \textbf{Metric Data Tables:} Basically this view provides a table that matches each UML model element analysed (table line) to it's value for each metric (table column);
\item \textbf{Histograms:} provides a graphical distribution  for each design metric;
\item \textbf{Design Comparison:} this view provides us a mean to compare the structural properties of two \textit{UML} designs. It is very useful to compare the same design in different iterations of the development, or to compare an alternative design to the current one.
\item \textbf{Rule Checker:} design rules and heuristics detect potencial problems in the UML design such as: 
	\begin{itemize}
	\item incomplete design such as unnamed classes, states without transitions;
	\item violation of naming conventions for classes, attributes, operations, packages;
	\item etc.
	\end{itemize}
\item \textbf{Catalog:} the catalog view provides us with the definitions of the metrics, design rules, and relation matrices for the current data set, and provides literature references and a glossary for them.
\end{itemize}
Not a view, but one of the most advanced features in this software is the possibility of defining Custom Design Metrics and Rules. The new metrics are defined in a XML file, with a very particular format, the \textit{SDMetricsML} (SDMetrics Markup Language). \\
The SDMetrics tool doesn't provide a direct notion of good/bad quality of the design model. Despite that, on the SDMetrics website we can find several indications of how to interpret each output. 
 
\subsubsection{Results}